A molecular dynamics study of the solvation of carbon dioxide and other compounds in the ionic liquids \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}

### 2018-09-14 ionic_liquids.tex
Simulation strategy for emin-NTf2

Another set of parameters for \ce{[emim][B(CN)_4]} was derived by Liu \textit{et al}. \cite{Liu_2014} in a study that investigates the dynamics of \ce{CO_2} and \ce{N_2} dissolved in this ionic liquid.
Parameters corresponding to the intramolecular and LJ interactions were extracted from GAFF \cite{Wang_2004}, while the atomic partial charges were computed using the RESP method \cite{Bayly_1993} with B3LYP/6-311+G(d) DFT calculations.
The net charges of the ions were manually adjusted to $\pm 0.8 \mathrm{e}$.

Finally, we also consider the parameterization which Weber and Kirchner \cite{Weber_2016} carried out for \ce{[emim][B(CN)_4]} based on the force field developed by P{\'{a}}dua and Canongia Lopes \cite{Canongia_Lopes_2006}.
As the potential parameters for the anion were specified neither in Ref.~\citenum{Weber_2016} nor in Ref.~\citenum{Canongia_Lopes_2006}, we obtained them by direct communication with the authors of Ref.~\citenum{Weber_2016}.
The atomic partial charges were derived using the RESP procedure with Hartree-Fock calculations at the \mbox{6-31++G\textsuperscript{**}} level of theory, resulting in net charges of $\pm 0.8 \mathrm{e}$.

@@

As a guide to the discussions we put forward in this section, in Fig~\ref{fig:atoms_id} we introduce a nomenclature for all atoms present in the ions \ce{[emim]^+}, \ce{[B(CN)_4]^-}, and \ce{[NTf_2]^-}.
Please note that, in the united-atom version of \ce{[emim]^+}, all hydrogen atoms ranging from H4 to H11 are incorporated to their corresponding carbon atoms.

@@

Using the software TRAVIS \cite{Brehm_2011}, we obtained combined distribution functions (CDFs) of the dihedral angles formed in \ce{[emim]^+} by atoms C1-N2-C5-C6 and atoms C2-N2-C5-C6.
These angles will be referred to as $\psi_1$ and $\psi_2$, respectively.
In Fig.~\ref{fig:dihedrals_emim}, we show the CDFs obtained for all force fields employed here for \ce{[emim][B(CN)_4]}.
The CDFs were generated from NPT-MD simulations of 250 ion pairs at $T = 298.0~\mathrm{K}$ and $P = 1.0~\mathrm{bar}$ using our standard setup (i.e. flexible ions with PPPM Ewald electrostatics).
Initial inspection reveals that, in all cases, $\psi_1$ and $\psi_2$ populate regions of the diagrams around the two straight lines given by
\begin{equation}
\label{eq:dihedral angles}
\psi_2 = \begin{cases}
\psi_1 + 180^\circ & \text{if} \quad \psi_1 \leq 180^\circ \\
\psi_1 - 180^\circ & \text{if} \quad \psi_1 > 180^\circ
\end{cases},
\end{equation}
both corresponding to a coplanar structure formed by atoms C1, C2, N2, and C5.
There are two such lines in each diagram because atom C6 can be located on either side of the plan.
One can note in Fig.~\ref{fig:dihedrals_emim} that the CDF peaks, which indicate the most likely dihedral angle combinations, lie either precisely above these lines or very close to them.
The largest distortions from coplanarity occur for the Batista \textit{et al}. \cite{Batista_2015} model, as observed by the peak locations in Fig.~\ref{fig:dihedrals_emim}(b).
Despite this, all rigid bodies considered here for \ce{[emim]^+} are built in such a way that atoms C[1-5], H[1-3], and N[1-2] all lie in a single plan.
Therefore, for specifying the torsion of bond N2-C5 it is enough to set a value for $\psi_1$ and then compute $\psi_2$ by means of Eq.~\eqref{eq:dihedral angles}.
All other structural features (bond lengths, bending angles, and hydrogen-involving dihedrals) are taken from the minimum-energy structure obtained for each force field.

@@

A study of dihedrals was also carried out for \ce{[emim][NTf_2]}.
The CDF of the dihedrals angles of C1-N2-C5-C6 and C2-N2-C5-C6, shown in Fig.~\ref{fig:die_ntf2}(a), resembles the results of Fig.~\ref{fig:dihedrals_emim}(d), but the most likely angle of C1-N2-C5-C6 is 0$^{\circ}$.
This similarity was expected as both models of \ce{[emim]^+} are based on the force field developed by Canongia-Lopes and P\'{a}dua~ \cite{Canongia_Lopes_2006}.
With regards to the anion \ce{[NTf_2]^-}, Deetlefs \textit{et al}. \cite{Deetlefs_2006} have experimentally determined the conformer population of the liquid-phase and observed that \ce{[NTf_2]^-} is mostly in its \textit{trans} conformation.
Fig.~\ref{fig:die_ntf2}(b), which corresponds to the CDF for the contiguous C1-S1-N1-S2 and S1-N1-S2-C2 chains, qualitatively verifies the observation of Deetlefs \textit{et al}. \cite{Deetlefs_2006}.
Given the considerable flexibility of \ce{[NTf_2]^-}, a more rigorous strategy for building the rigid bodies is required.
Thus, the simulations of \ce{[emim][NTf_2]} were performed considering flexible ions and the DSF method, which results in a computed density of 1.495 $\pm$ 0.009 g/cm$^3$ with an error of 1.12\% with respect to the experimental value~\cite{Tokuda_2005}.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{Ludwig_anion}%
\caption{Combined distribution function of the dihedral angles of the contiguous chains C1-S1-N1-S2 and S1-N1-S2-C2 of anion \ce{[NTf_2]^-}, obtained from NPT-MD simulations of 250 ion pairs of \ce{[emim][NTf_2]} using the model of K\"{o}ddermann \textit{et al}. \cite{K_ddermann_2007}.}
\label{fig:die_ntf2}
\end{figure}

>>>>

Another set of parameters for \ce{[emim][B(CN)_4]} was derived by Liu \textit{et al}. \cite{Liu_2014} in a study that investigates the dynamics of \ce{CO_2} and \ce{N_2} dissolved in this ionic liquid.
Parameters corresponding to the intramolecular and LJ interactions were extracted from GAFF \cite{Wang_2004}, while the atomic partial charges were computed using the RESP method \cite{Bayly_1993} with B3LYP/6-311+G(d) DFT calculations.
The net charges of the ions were manually adjusted to $\pm 0.8 ~\mathrm{e}$.

Finally, we also consider the parameterization which Weber and Kirchner \cite{Weber_2016} carried out for \ce{[emim][B(CN)_4]} based on the force field developed by P{\'{a}}dua and Canongia Lopes \cite{Canongia_Lopes_2006}.
As the potential parameters for the anion were specified neither in Ref.~\citenum{Weber_2016} nor in Ref.~\citenum{Canongia_Lopes_2006}, we obtained them by direct communication with the authors of Ref.~\citenum{Weber_2016}.
The atomic partial charges were derived using the RESP procedure with Hartree-Fock calculations at the \mbox{6-31++G\textsuperscript{**}} level of theory, resulting in net charges of $\pm 0.8 ~\mathrm{e}$.

@@

As a guide to the discussions in this section, in Fig~\ref{fig:atoms_id} we introduce a nomenclature for all atoms present in the ions \ce{[emim]^+}, \ce{[B(CN)_4]^-}, and \ce{[NTf_2]^-}.
Please note that, in the united-atom version of \ce{[emim]^+}, all hydrogen atoms ranging from H4 to H11 are incorporated to their corresponding carbon atoms.

@@

Using the software TRAVIS \cite{Brehm_2011}, we obtained combined distribution functions (CDFs) of the dihedral angles formed in \ce{[emim]^+} by atoms C1-N2-C5-C6 and atoms C2-N2-C5-C6.
These angles will be referred to as $\psi_1$ and $\psi_2$, respectively.
In Fig.~\ref{fig:dihedrals_emim}, we show the CDFs obtained for all force fields employed here for \ce{[emim][B(CN)_4]}.
All CDFs shown in this section were generated from NPT-MD simulations of 250 ion pairs at $T = 298.0~\mathrm{K}$ and $P = 1.0~\mathrm{bar}$ using our standard setup (i.e. flexible ions with PPPM Ewald electrostatics).
Initial inspection reveals that, in all cases, $\psi_1$ and $\psi_2$ populate regions of the diagrams around the two straight lines given by
\begin{equation}
\label{eq:dihedral angles}
\psi_2 = \begin{cases}
\psi_1 + 180^\circ & \text{if} \quad \psi_1 \leq 180^\circ \\
\psi_1 - 180^\circ & \text{if} \quad \psi_1 > 180^\circ
\end{cases},
\end{equation}
both corresponding to a coplanar structure formed by atoms C1, C2, N2, and C5.
There are two such lines in each diagram because atom C6 can be located on either side of the plan.
One can note in Fig.~\ref{fig:dihedrals_emim} that the CDF peaks, which indicate the most likely dihedral angle combinations, lie either precisely above these lines or very close to them.
The largest distortions from coplanarity occur for the Batista \textit{et al}. \cite{Batista_2015} model, as observed by the peak locations in Fig.~\ref{fig:dihedrals_emim}(b).
Despite this, all rigid bodies considered here for \ce{[emim]^+} are built in such a way that atoms C[1-5], H[1-3], and N[1-2] all lie in a single plan.
Therefore, for specifying the torsion of bond N2-C5 it is enough to set a value for $\psi_1$ and then compute $\psi_2$ by means of Eq.~\eqref{eq:dihedral angles}.
All other structural features (bond lengths, bending angles, and hydrogen-involving dihedrals) are taken from the minimum-energy structure obtained for each force field.

@@

A study of conformations was also carried out for \ce{[emim][NTf_2]} using the model of K\"{o}ddermann \textit{et al}. \cite{K_ddermann_2007}.
A CDF obtained for the dihedral angles of sequences C1-N2-C5-C6 and C2-N2-C5-C6 in \ce{[emim]^+}, which is made available in the Supplementary Material, resembles the one in Fig.~\ref{fig:dihedrals_emim}(d).
The resemblance was expected because the intramolecular parameters employed for \ce{[emim]^+} are identical in these two cases, since the corresponding models are both based on the force field developed by Canongia-Lopes and P\'{a}dua \cite{Canongia_Lopes_2006}.
However, the most likely value of $\psi_1$ is shifted to $0^{\circ}$ in the case of \ce{[emim][NTf_2]}.
This highlights the fact that altering atomic charges can induce conformational changes and should, therefore, be accompanied by a revision of torsion parameters in some cases.
With regard to the anion \ce{[NTf_2]^-}, Deetlefs \textit{et al}. \cite{Deetlefs_2006} experimentally determined the conformer population of the liquid-phase and observed that \ce{[NTf_2]^-} is mostly found in its \textit{trans} conformation.
In Fig.~\ref{fig:die_ntf2}, we present the CDF obtained for the contiguous sequences C1-S1-N1-S2 and S1-N1-S2-C2 in \ce{[NTf_2]^-}, which qualitatively confirms the observation of Deetlefs \textit{et al}. \cite{Deetlefs_2006}.
This CDF also makes clear that \ce{[NTf_2]^-} exhibits a considerable degree of flexibility.
Due to this observed behaviors of both ions, all simulations of \ce{[emim][NTf_2]} in the present work are performed considering flexible cations and anions, as well as DSF electrostatics.
This approach results in a mean density of $1.495 \pm 0.009 ~\mathrm{g/cm^3}$ at $T = 298.0~\mathrm{K}$ and $P = 1.0~\mathrm{bar}$.
The relative error of this result with respect to the experimental value obtained by Tokuda \textit{et al}. \cite{Tokuda_2005} is only 1.12\%.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{Ludwig_anion}
\caption{Combined distribution function of the dihedral angles of the contiguous chains C1-S1-N1-S2 and S1-N1-S2-C2 of anion \ce{[NTf_2]^-}, obtained from NPT-MD simulations of 250 ion pairs of \ce{[emim][NTf_2]} using the model of K\"{o}ddermann \textit{et al}. \cite{K_ddermann_2007}.}
\label{fig:die_ntf2}
\end{figure}
